---
title: "ml assignment 1"
author: "nemin dholakia"
date: "9/10/2021"
output: html_document
 df print:paged
 
 
```{r setup , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Q1: Data source
Our dataset **Overwise statistics without superover** has been extracted from **KAGGLE** , is a collection of database and datasets used by Machine Learning Community for the analysis of Machine Learning algorithms.

###Q2: Importing Overwise statistics without superover dataset to R Studio

```{r }
setwd("C:/Users/User/Downloads/archive (1)")
OS_WS <- read.csv("Overwise statistics without super over.csv")
head(OS_WS)

```
###Q3: Descriptive analysis of OWSW Dataset
Inspecting dataset
```{r message=FALSE }
summary(OS_WS)

```
**Arithmetic Mean: ** AM of a set of observation is defined as their sum divided number of observations

```{r }
#Average runs 
mean(OS_WS$runs)
```
**Median:** is the middle value of the observation.

```{r }
#Median of wickets
median(sort(OS_WS$wickets))
```
**Mode :** refers to the value which occurs to the maximum frequency.

```{r }
sort(table(OS_WS$winner),decreasing=TRUE)
```

###Q4: Transformation

```{r }
#Applying log transformation on Wicket column
run <- log10(OS_WS$runs)
##displaying first 10 values
run[1:10]

```
###Q5:Visualisation
**Histogram Plot**

```{r message=FALSE, warning= FALSE }
library(ggplot2)
ggplot(OS_WS,aes(x= wickets))+geom_histogram()+ggtitle('Histogram')

```
**Scatter Plot**

```{r }
ggplot(OS_WS, aes(x= runs, y=wickets))+geom_point()+ggtitle('Scatter Plot')
```